---
title: A tour of the Hector R interface
author: Alexey Shiklomanov
output:
  xaringan::moon_reader
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)

# Invisible versions of Hector functions so they don't pollute the output
run <- function(...) invisible(hector::run(...))
reset <- function(...) invisible(hector::reset(...))
```

# Install directly from GitHub

```{r install, eval = FALSE}
Sys.setenv(
  BOOSTROOT = "/path/to/boost/include",
  BOOSTLIB = "/path/to/boost/lib"
)
devtools::install_github("jgcri/hector")
```

---

# Lightning introduction

```{r}
library(hector)
ini_file <- system.file(
  "input/hector_rcp45.ini",
  package = "hector"
)
core <- newcore(ini_file)
run(core)
results <- fetchvars(core, dates = 2000:2100)
head(results)
```

---

### Lightning introduction (continued)

```{r, fig.width = 6, fig.height = 3.3, fig.align = "center"}
results$variable_unit <- with(results, {
  sprintf("%s (%s)", variable, units) 
})
library(ggplot2)
ggplot(results) +
  aes(x = year, y = value) +
  geom_line() +
  facet_wrap(~variable_unit, scales = "free_y")
```

---

## Retrieving parameters

```{r}
fetchvars(core, dates = NA, vars = c(BETA(), Q10_RH(), ECS()))
```

`BETA` is the CO<sub>2</sub> fertilization factor.

`Q10_RH` is the sensitivity of heterotropihc respiration to temperature.

`ECS` is the equilibrium climate sensitivity.

...and more! See `?hector::parameters`.

---

## Setting parameters

Increase `BETA` to 0.4 and re-run the RCP 4.5 scenario.

```{r}
setvar(
  core,
  dates = NA,
  var = BETA(),
  value = 0.4,
  unit = "(unitless)"
)

reset(core)
run(core)
new_results <- fetchvars(core, 2000:2100)
```

--

Combine with previous results.

```{r}
new_results$beta <- 0.4 
new_results$variable_unit <- with(new_results, {
  sprintf("%s (%s)", variable, units)
})
results$beta <- 0.36
both_results <- rbind(new_results, results)
```

---

```{r, fig.align = "center"}
ggplot(both_results) +
  aes(x = year, y = value, color = factor(beta)) +
  geom_line() +
  facet_wrap(~variable_unit, scales = "free_y") +
  guides(color = guide_legend(title = expression(beta)))
```

---

## Case study: Parameter sensitivity analysis

```{r}
run_with_beta <- function(value) {
  setvar(core, NA, BETA(), value, "(unitless)")
  reset(core)
  run(core)
  result <- fetchvars(core, 2000:2300)
  result[["beta"]] <- value
  result[["variable_unit"]] <- with(result, {
    sprintf("%s (%s)", variable, units)
  })
  result
}

mapped <- Map(run_with_beta, seq(0, 1, 0.05))
sensitivity_beta <- Reduce(rbind, mapped)
```

---

```{r, fig.align = "center", fig.with = 6, fig.height = 4.8}
ggplot(sensitivity_beta) +
  aes(x = year, y = value, color = beta, group = beta) +
  geom_line() +
  facet_wrap(~variable_unit, scales = "free_y") +
  guides(color = guide_colorbar(title = expression(beta))) +
  scale_color_viridis_c()
```
